---
title: "gallup_world_relative_income_analysis_grace"
author: "Yujia Bai (Grace)"
date: '2022-04-22'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#file path: "G:/My Drive/projects/income_comparisons/code/gallup_world/gallup_world_relative_income_plots_grace.Rmd
```

## Plotting
```{r}
#rename variables
country_level_models <- 
  country_level_models %>% 
  dplyr::rename(comp_income_effect = `scale(mi_value)`)

final_df <- 
  final_df %>% 
  dplyr::rename(comp_income_effect = `scale(mi_value)`)
```


### Stargazer plot for regression analysis
#### demographic median income
```{r}
#done in regression_analysis_gallup_world_US_grace
```


### Gallup World model fits
- summary stats

- histogram for coefs

#### demographic median income
```{r}
#function for plotting histograms
# plot_data_column = function (data, column) {
#     ggplot(data, aes_string(x = column)) +
#         geom_histogram(fill = "darkgrey", bins=50) +
#         xlab(column)
# }


#plot comparison income coef
plot_1 <- country_level_models %>% 
  filter(mi_var == "demographic_median_income") %>% 
  ggplot(aes(comp_income_effect)) +
  geom_histogram(bins = 50) +
  geom_vline(xintercept = -0.038) +
  xlab("Relative income effect") +
  ggtitle("Demographic median comparison income")+
  theme_bw()

 ggsave("plot_1.png")

#plot other coefs
# dem_med_income_plot <- lapply(colnames(dem_med_income), 
#                               plot_data_column, 
#                               data = dem_med_income)
# 
# dem_med_income_plot
```


```{r}
dem_med_income <- filter(country_level_models, 
                         mi_var == "demographic_median_income") %>%
  select(c(1,4:20)) %>% na.omit %>% as.data.frame()

#summary stats
summ1 <- lapply(dem_med_income, function(x) {
   paste0( "[",
     # " (95% CI: ",
   paste(round(sort(mean(x) + c(1.96, -1.96) * sd(x)/sqrt(length(x))), 3),
         collapse = ", "), "]")
 } ) %>%
  rbind(summarise_all(dem_med_income, sd)) %>%
  rbind(summarise_all(dem_med_income, mean)) %>%
  rbind(summarise_all(dem_med_income, median)) %>%
  as.data.frame()

rownames(summ1) <- c("95%CI","SD","Mean","Median")
summ1
write_csv(summ1, "Gallup World_dem_median_models_summary.csv")

#order relative income effects
dem_med_income[order(dem_med_income$comp_income_effect),] %>% select(c(18),everything())
```


#### demographic rf income
```{r}
## for demographic rf income

#plot comparison income coef
plot_2 <- country_level_models %>% 
  filter(mi_var == "demographic_rf_income") %>% 
  ggplot(aes(comp_income_effect)) +
  geom_histogram(bins = 50) +
  geom_vline(xintercept = -0.0142) +
  xlab("Relative income effect") +
  ggtitle("Demographic random forest predicted income")+
  theme_bw()

 ggsave("plot_2.png")
 

#plot other coefs
# dem_rf_income_plot <- lapply(colnames(dem_rf_income), 
#                               plot_data_column, 
#                               data = dem_rf_income)
# 
# dem_med_income_plot
 
 
#summary stats
dem_rf_income <- filter(country_level_models, mi_var == "demographic_rf_income") %>%
  select(c(1,4:20)) %>% na.omit() %>% as.data.frame()

summ2 <- lapply(dem_rf_income, function(x) {
   paste0( "[",
     # " (95% CI: ",
   paste(round(sort(mean(x) + c(1.96, -1.96) * sd(x)/sqrt(length(x))), 3),
         collapse = ", "), "]")
 } ) %>%
  rbind(summarise_all(dem_rf_income, sd)) %>%
  rbind(summarise_all(dem_rf_income, mean)) %>%
  rbind(summarise_all(dem_rf_income, median)) %>%
  as.data.frame()

rownames(summ2) <- c("95%CI","SD","Mean","Median")
summ2
write_csv(summ2, "Gallup World_dem_rf_models_summary.csv")

dem_rf_income[order(dem_rf_income$comp_income_effect),] %>% select(c(18),everything())
```

### Results correlating comparison income coefficient with secondary data variables

The standardized coefficient (correlation coefficient)

95% confidence interval around the coefficient

The t-statistic

P-value

--------------------------------------------------------------------------------------------------
Each row will showcase the relationship between comparison income and one variable from the secondary datasets:

GDP - calculate the average from 2003 to 2007
Mean / median income at the country-level
Gini Index
Change in GDP average of 1988 to 1992 compared to 2003 to 2007
6 Hoftstede variables
Variables from the WVS

#### demographic median income
vars: scale(income_usd_log); gini; hof; all wvs
```{r}
final_df
```

##### create dataframe
```{r}
#use new gdp and gini data
#create dataframe
dem_med_income0 <- filter(final_df, mi_var == "demographic_median_income")

dem_med_income <- select(dem_med_income0, c(1, 20, 21:26, 41,46, 47:60)) %>%
  mutate(country = as.factor(country)) %>%
  mutate_if(is.character, as.double)

head(dem_med_income)
```

#####summary stats
```{r}
#summary stats
#cannot run with country in df
lapply(dem_med_income, function(x) {
   paste0( "[",
     # " (95% CI: ",
   paste(round(sort(mean(x) + c(1.96, -1.96) * sd(x)/sqrt(length(x))), 2),
         collapse = ", "), "]")
 } ) %>%
  rbind(summarise_all(dem_med_income, sd)) %>%
  rbind(summarise_all(dem_med_income, mean)) %>%
  rbind(summarise_all(dem_med_income, median))


#calculate avg GDP from 2003 - 2007: produces NAs
# dem_med_income$gdp_avg <- mean(sum(dem_med_income[,24:28]))
#if lose data > 1/3: make interval wider: 2006~2018
```


```{r}
#correlation with gdp and gini were not significant
library(Hmisc)
corr1 <- rcorr(as.matrix(dem_med_income),type="pearson") %>% as.list()
corr1 <- data.frame(corr1$P)
corr1 # Significant: idv;mas;Q150


dem_med_income <- dem_med_income[,c("comp_income_effect","idv","mas","Q150")]
corr_dem_med_income <- rcorr(as.matrix(dem_med_income),type="pearson") %>% as.list()
corr_dem_med_income <- corr_dem_med_income$P%>%as.data.frame()


corr_dem_med_income <- corr_dem_med_income$comp_income_effect%>%as.data.frame()
rownames(corr_dem_med_income)<-c("comp_income_effect","idv","mas","Q150")
colnames(corr_dem_med_income) <- "comp_income_effect_pval"
corr_dem_med_income

```


#### Scatterplots
```{r}
## Scatterplots: for age_scale and gender2 and Q150
# source("G:/My Drive/projects/income_comparisons/code/gallup_world/fdr.R")
# fdr_cut(corr_dem_med_income$comp_income_effect_pval, 0.01, plotit = T) #cutoff is 0.002

# lm1 <- lm(dem_med_income$comp_income_effect ~ dem_med_income$idv)
# plot(dem_med_income$comp_income_effect ~ dem_med_income$idv)
# abline(lm1)
# 
# lm2 <- lm(dem_med_income$comp_income_effect ~ dem_med_income$mas)
# plot(dem_med_income$comp_income_effect ~ dem_med_income$mas)
# abline(lm2)
# 
# lm_3 <- lm(dem_med_income$comp_income_effect ~ dem_med_income$Q150)
# plot(dem_med_income$comp_income_effect ~ dem_med_income$Q150)
# abline(lm_3)

#use ggplot
dem_med_idv <- ggplot(dem_med_income,aes(idv, comp_income_effect, label = country))+
  geom_text() +
  xlab("Individualism vs. Collectivism") +
  ylab("Relative income effect") +
  stat_smooth(method="lm", se=T)
dem_med_idv
ggsave("dem_med_idv.png")

dem_med_mas <- ggplot(dem_med_income,aes(mas, comp_income_effect, label = country))+
  geom_text() +
  xlab("Masculinity vs. Femininity") +
  ylab("Relative income effect") +
  stat_smooth(method="lm", se=T)
dem_med_mas
ggsave("dem_med_mas.png")

dem_med_free_secure <- ggplot(dem_med_income,aes(Q150, comp_income_effect, label = country))+
  geom_text() +
  xlab("Freedom vs. Security") +
  ylab("Relative income effect") +
  stat_smooth(method="lm", se=T)
  #theme(aspect.ratio = 3/4)
dem_med_free_secure
ggsave("dem_med_free_secure.png", width = 12, height = 4)


library(gridExtra)
gridExtra::grid.arrange(dem_med_idv,dem_med_mas,dem_med_free_secure, nrow = 1)
```
Significant:
age;educ2;employmen4&5;gender2;idv;mas;Q150

#### demographic rf income
```{r}
dem_rf_income <- filter(final_df, 
                         mi_var == "demographic_rf_income") %>% 
  select(c(1, 20, 21:26, 41,46, 47:60)) %>%
  mutate(country = as.factor(country)) %>%
  mutate_if(is.character, as.double)

head(dem_rf_income)

#summary stats
#cannot run with country in df; same thing in dem_med_income
lapply(select(dem_rf_income), function(x) {
   paste0( "[",
     # " (95% CI: ",
   paste(round(sort(mean(x) + c(1.96, -1.96) * sd(x)/sqrt(length(x))), 2),
         collapse = ", "), "]")
 } ) %>%
  rbind(summarise_all(dem_rf_income, sd)) %>%
  rbind(summarise_all(dem_rf_income, mean)) %>%
  rbind(summarise_all(dem_rf_income, median))
```


```{r}
library(Hmisc)
corr2 <- rcorr(as.matrix(select(dem_rf_income,-c(1))),type="pearson") %>% as.list()
corr2 <- data.frame(corr2$P)
corr2

#significant vars
dem_rf_income <- dem_rf_income[,c("comp_income_effect","pdi","idv","ivr","gdp","Q50")]


corr_dem_rf_income <- rcorr(as.matrix(dem_rf_income),type="pearson") %>% as.list()
corr_dem_rf_income <- corr_dem_rf_income$P %>% as.data.frame()
corr_dem_rf_income <- corr_dem_rf_income$comp_income_effect%>%as.data.frame()

rownames(corr_dem_rf_income) <-c("comp_income_effect","pdi","idv","ivr","gdp","Q50")

colnames(corr_dem_rf_income) <- "comp_income_effect_pval"
corr_dem_rf_income
```

#### Scatterplots
```{r}
## Scatterplots: for age_scale, gender2, and  marital_status2
#fdr_cut(corr_dem_rf_income$comp_income_effect_pval, 0.01, plotit = T) #cutoff is 0.001

lm4 <- lm(dem_rf_income$comp_income_effect ~ dem_rf_income$ivr)
plot(dem_rf_income$comp_income_effect ~ dem_rf_income$ivr)
abline(lm_age_scale)

lm5 <- lm(dem_rf_income$comp_income_effect ~ dem_rf_income$pdi)
plot(dem_rf_income$comp_income_effect ~ dem_rf_income$pdi)
abline(lm5)

lm6 <-  lm(dem_rf_income$comp_income_effect ~ dem_rf_income$idv)
plot(dem_rf_income$comp_income_effect ~ dem_rf_income$idv)
abline(lm6)


#use ggplot
dem_rf_ivr <- ggplot(dem_rf_income,aes(ivr, comp_income_effect, label = country))+
  geom_text() +
  xlab("Indulgence") +
  ylab("Relative income effect") +
  stat_smooth(method="lm", se=T)
dem_rf_ivr
ggsave("dem_rf_ivr.png")


dem_rf_pdi <- ggplot(dem_rf_income,aes(pdi, comp_income_effect, label = country))+
  geom_text() +
  xlab("Power Distance") +
  ylab("Relative income effect") +
  stat_smooth(method="lm", se=T)
dem_rf_pdi
ggsave("dem_rf_pdi.png")


dem_rf_idv <- ggplot(dem_rf_income,aes(idv, comp_income_effect, label = country))+
  geom_text() +
  xlab("Individualism vs. Collectivism") +
  ylab("Relative income effect") +
  stat_smooth(method="lm", se=T)
dem_rf_idv
ggsave("dem_rf_idv.png")


dem_rf_gdp <- ggplot(dem_rf_income,aes(log(gdp), comp_income_effect, label = country))+
  geom_text() +
  xlab("log(Averge GDP from 2006 to 2018)") +
  ylab("Relative income effect") +
  stat_smooth(method="lm", se=T)
dem_rf_gdp
ggsave("dem_rf_gdp.png")

library(gridExtra)
gridExtra::grid.arrange(dem_rf_idv, dem_rf_pdi, dem_rf_ivr, dem_rf_gdp, nrow=2)
```
